---
title: "Case_Study_2"
author: "Ben Tanaka"
date: "7/29/2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, echo=FALSE, include=FALSE, warning=FALSE,message =FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,message =FALSE)
library(xml2)
library(rvest)
library(stringr)
library(tibble)
library(data.table)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(ggplot2)
library(bizdays)
library(MTS)
library(forecast)
library(smooth)
library(fpp2)
library(lubridate)
library(dygraphs)
library(xts)
library(readxl)
library(tidyverse)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(GGally)
library(psych)
library(scales)
library(cowplot)

setwd("/users/ben/documents/r/Case_Study_2")
getwd()
```

# DDSAnalytics:  Predicting Employee Turnover

## Executive Summary
DDSAnalytics looks to be the Data Science partner of choice in which it looks to attract and retain its employee base.  Employee retention will allow the company to further deepen and cement its place as a competitive data science organization.  Currently there we have information about the DDSAnaltics employees that contains their demographic as well as employee metadata that will allow us to identify the trends and factors that could potentially lead to higher attrition rates.

An Exploratory Data Analysis was conducted on the dataset which invovled cleaning the data, ensuring the data has limited missing values, the data was transformed to provide the correct elements for analysis, conducting correlation analysis, and graphing the data to identify trends.

The study found that employees that the following factors had higher rates of attrition; Below College Education, singe marital status employees, employees with less than 10 years working experience, employees who live more than 10 miles away from work, employees who are younger than 30,  employees who make less than $5k a month, low job involvement, low environment satisfaction, sales department and sales representatives, less than 5 years with the manager and a current role.

Additional work should be done to deep dive into these factors to identify further trends within these smaller segements.  Once an analysis is completed within these smaller segments a Random Forest for antoher type of classification model should be generated to identify current employees who could be at risk to leave the company.

```{r ETL}
df_emp_ren <- read_excel("CaseStudy2-data.xlsx")
names(df_emp_ren)

df_emp_ren$Attrition_Lvl <- factor(df_emp_ren$Attrition)
str(df_emp_ren)
summary(df_emp_ren)

```
## Transform Factors into actual results

``` {r Transform Factors}
df_emp_ren <- df_emp_ren %>%
  mutate(Education = as.factor(if_else(Education == 1,"Below College", if_else(Education == 2, "College", if_else(Education == 3, "Bachelor", if_else(Education == 4, "Master","Doctor")))))
         ,EnvironmentSatisfaction = as.factor(if_else(EnvironmentSatisfaction == 1,"Low",if_else(EnvironmentSatisfaction == 2, "Medium", if_else(EnvironmentSatisfaction == 3, "High", "Very High"))))
         ,JobInvolvement = as.factor(if_else(JobInvolvement == 1,"Low",if_else(JobInvolvement == 2, "Medium",if_else(JobInvolvement == 3, "High", "Very High"))))
         ,JobSatisfaction = as.factor(if_else(JobSatisfaction == 1, "Low",if_else(JobSatisfaction == 2, "Medium",if_else(JobSatisfaction == 3, "High","Very High"))))
         ,PerformanceRating = as.factor(if_else(PerformanceRating == 1, "Low",if_else(PerformanceRating == 2, "Good", if_else(PerformanceRating == 3, "Excellent", "Outstanding"))))
         ,RelationshipSatisfaction = as.factor(if_else(RelationshipSatisfaction == 1, "Low",if_else(RelationshipSatisfaction == 2, "Medium", if_else(RelationshipSatisfaction == 3, "High", "Very High"))))
         ,WorkLifeBalance = as.factor(if_else(WorkLifeBalance == 1, "Bad",if_else(WorkLifeBalance == 2, "Good", if_else(WorkLifeBalance == 3, "Better", "Best"))))
         ,JobLevel = as.factor(JobLevel)
         )
```

## Employee Personal Demographics
``` {r Employee_Demographics}

p1 <- ggplot(df_emp_ren, aes(x=EducationField ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="EducationField") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Education Field")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p2 <- df_emp_ren %>%
  group_by(EducationField) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = EducationField, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Education Field") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p3 <- ggplot(df_emp_ren, aes(x=Education ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count",  size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="Attrition_Lvl") +
  theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
  theme(legend.position = "none") +
  ggtitle("Attrition Counts - Education")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p4 <- df_emp_ren %>%
  group_by(Education) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = Education, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Education") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))


grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)


EmpPerDem2 <- df_emp_ren[c("NumCompaniesWorked","TotalWorkingYears", "DistanceFromHome" ,"Attrition_Lvl")]
ggpairs(EmpPerDem2)

p1 <- df_emp_ren %>%
  ggplot(aes(x = NumCompaniesWorked, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Number Companies Worked")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- df_emp_ren %>%
  ggplot(aes(x = TotalWorkingYears, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Total Working Years")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- df_emp_ren %>%
  ggplot(aes(x = DistanceFromHome, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Distance From Home")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

grid.arrange(p1, p2, p3, nrow = 2, ncol = 2)


p1 <- ggplot(df_emp_ren, aes(x=MaritalStatus ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="MaritalStatus") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Marital Status")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p2 <- df_emp_ren %>%
  group_by(MaritalStatus) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = MaritalStatus, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - MaritalStatus") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))


p3 <- ggplot(df_emp_ren, aes(x=Gender ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="Gender") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Gender")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p4 <- df_emp_ren %>%
  group_by(Gender) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = Gender, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Gender") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))


grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)

EmpPerDem3 <- df_emp_ren[c("Age","Attrition_Lvl")]
ggpairs(EmpPerDem3)

p3 <- df_emp_ren %>%
  ggplot(aes(x = Age, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Age")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3
```

### Employee Personal Demographics:  Insights
* Insights:
    1.  Employees with an education field in Life Science and Medical tend to have higher counts of attrition however those who have a background in Human Resources and/or a Technical degree tend to have higher attrition rates.
      + Hypothesis: Human Resource employees drive the most attrition rates.
    2.  Employees with a Bachelor & Masters degree have higher counts of attrition, whereas, those with a below college degree tend to have higher rates of attrition.
      + Hypothesis:  Employees with a human resources education field and a below college education will have higher rates of attrition.
    3.  The more companies a person works for the higher the attrtion level.  
    4.  There was a higher amount rate of employees who leave when they have less working years.  
    5.  The Distance from home was can be used as a factor for higher attrition rates.  As the farther away the employee is the more likely they are to leave.
    + Hypothesis:  Greater distance from work will drive higher attrition.
    6.  People who are single tend to have higher attrition rates
      +  Hypothesis:  Employess who have not been in a commited relationship could leave the company.  Singles nights at work could help drive lower attrition rates.  
    7.  Males tend to have a higher attrition rate although not to a significant degreee
      + Hypothesis:  Gender does not contribute to higher attrition rate.
    8.  Employees who are younger tend to have higher attrition rates.
      + Hypothesis:  Younger employees are more likely to leave the company
      


## Employee Pay Rates
``` {r Employee Pay Rates}
EmpPayRte1 <- df_emp_ren[c("HourlyRate", "DailyRate", "MonthlyRate","MonthlyIncome", "PercentSalaryHike","Attrition_Lvl")]
ggpairs(EmpPayRte1)

p1 <- df_emp_ren %>%
  ggplot(aes(x = HourlyRate, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Hourly Rate")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- df_emp_ren %>%
  ggplot(aes(x = DailyRate, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Daily Rate")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- df_emp_ren %>%
  ggplot(aes(x = MonthlyRate, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Monthly Rate")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p4 <- df_emp_ren %>%
  ggplot(aes(x = MonthlyIncome, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Monthly Income")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p5 <- df_emp_ren %>%
  ggplot(aes(x = PercentSalaryHike, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Percent Salary Hike")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

grid.arrange(p1, p2, p3, p4, p5, nrow = 3, ncol = 3)

EmpPayRte2 <- df_emp_ren[c("StockOptionLevel","OverTime","Attrition_Lvl")]
ggpairs(EmpPayRte2)

p1 <- df_emp_ren %>%
  ggplot(aes(x = StockOptionLevel, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Stock Option Level")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- df_emp_ren %>%
  ggplot(aes(x = OverTime, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Over Time")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

### Employee Pay Rates:  Insights
* Insights:
    1.  Lower hourly & daily rates did experience higer attrition rates when they pay was lower.  This also impacted the Monthly Income in which basically, those employees who make less tend to leave.
      + Hypothesis: Employees with lower monthly incomes and daily/hourly rates tend to have higher attrition rates
    2.  The Month rates did have a small spike as the lower incomes however it was not conclusive it will be able to help as a identifying factor for attrition.
    3.  Employees with lower stock options have a higher rate of attrition.
    +  Hypothesis:  Employees with lower stock options will tend to leave the company.  Coupling stock option values as well as Monthly income will help generate a wholistic approach to work compensation
    4.  Overtime did show a slight indication of attrition however, it will generate higher monthly income which was shown to drive lower attrition.  
    +  Hypothesis:  Employees who work excessice overtime will tend to leave the company if it doesn't generate enough of a monthly income.

## Employee Work Profile Demographics
```{r Employee Work Profile Demographics}
EmpPflDem1 <- df_emp_ren[c("YearsWithCurrManager","YearsSinceLastPromotion","YearsInCurrentRole","YearsAtCompany","PerformanceRating", "Attrition_Lvl")]
ggpairs(EmpPflDem1)

p1 <- df_emp_ren %>%
  ggplot(aes(x = YearsWithCurrManager, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years With Current Manager")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- df_emp_ren %>%
  ggplot(aes(x = YearsSinceLastPromotion, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years Since Last Promotion")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- df_emp_ren %>%
  ggplot(aes(x = YearsInCurrentRole, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years In Current Role")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p4 <- df_emp_ren %>%
  ggplot(aes(x = YearsAtCompany, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years At Company")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p5 <- ggplot(df_emp_ren, aes(x=PerformanceRating ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="PerformanceRating") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Performance Rating")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p6 <- df_emp_ren %>%
  group_by(PerformanceRating) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = PerformanceRating, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Performance Rating") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
grid.arrange(p5,p6, nrow = 1, ncol = 2)

EmpPflDem2 <- df_emp_ren[c("WorkLifeBalance","RelationshipSatisfaction","JobSatisfaction","JobInvolvement","EnvironmentSatisfaction", "Attrition_Lvl")]
ggpairs(EmpPflDem2)

p1 <- ggplot(df_emp_ren, aes(x=WorkLifeBalance ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="WorkLifeBalance") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Work Life Balance")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p2 <- df_emp_ren %>%
  group_by(WorkLifeBalance) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = WorkLifeBalance, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Work life Balance") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p3 <- ggplot(df_emp_ren, aes(x=RelationshipSatisfaction ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="RelationshipSatisfaction") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Relationship Satisfaction")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p4 <- df_emp_ren %>%
  group_by(RelationshipSatisfaction) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = RelationshipSatisfaction, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Relationship Satisfaction") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p5 <- ggplot(df_emp_ren, aes(x=JobSatisfaction ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="JobSatisfaction") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Job Satisfaction")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p6 <- df_emp_ren %>%
  group_by(JobSatisfaction) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobSatisfaction, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Job Satisfaction") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p7 <- ggplot(df_emp_ren, aes(x=JobInvolvement ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="JobInvolvement") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Job Involement")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p8 <- df_emp_ren %>%
  group_by(JobInvolvement) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobInvolvement, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Job Involvement") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p9 <- ggplot(df_emp_ren, aes(x=EnvironmentSatisfaction ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="EnvironmentSatisfaction") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Environment Satisfaction")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p10 <- df_emp_ren %>%
  group_by(EnvironmentSatisfaction) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = EnvironmentSatisfaction, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Environment Satisfaction") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

grid.arrange(p1, p2, p3, p4,  nrow = 2, ncol = 2)
grid.arrange(p5, p6, p7, p8,  nrow = 2, ncol = 2)
grid.arrange( p9, p10, nrow = 1, ncol = 2)

EmpPflDem3 <- df_emp_ren[c("TrainingTimesLastYear","StandardHours","JobRole","JobLevel","Department","BusinessTravel", "Attrition_Lvl")]
ggpairs(EmpPflDem3)

p1 <- df_emp_ren %>%
  ggplot(aes(x = TrainingTimesLastYear, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Training Times Last Year")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- df_emp_ren %>%
  ggplot(aes(x = StandardHours, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Standard Hours")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- ggplot(df_emp_ren, aes(x=JobRole ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="JobRole") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Job Role")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p4 <- df_emp_ren %>%
  group_by(JobRole) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobRole, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Job Role") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p5 <- ggplot(df_emp_ren, aes(x=Department ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="Department") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Department")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p6 <- df_emp_ren %>%
  group_by(Department) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = Department, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Department") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 100))

p7 <- ggplot(df_emp_ren, aes(x=BusinessTravel ,  group=Attrition)) +
geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
geom_text(aes( label = scales::percent(..prop..),
y= ..prop.. ), stat= "count", size = 2.5, position=position_dodge(width=0.2)) +
labs(y = "Percent", fill="BusinessTravel") +
theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +
theme(legend.position = "none") +
ggtitle("Attrition Counts - Business Travel")+
facet_grid(~Attrition) +
scale_y_continuous(labels = scales::percent)

p8 <- df_emp_ren %>%
  group_by(BusinessTravel) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = BusinessTravel, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "coral3") + ggtitle("Attrition Rate - Business Travel") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0,100))

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
grid.arrange( p5,p6,p7,p8, nrow = 2, ncol = 2)

EmpPflDem4 <- df_emp_ren[c("YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion","YearsWithCurrManager", "Attrition_Lvl")]
ggpairs(EmpPflDem4)

p1 <- df_emp_ren %>%
  ggplot(aes(x = YearsAtCompany, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years At Company")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- df_emp_ren %>%
  ggplot(aes(x = YearsInCurrentRole, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years In Current Role")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- df_emp_ren %>%
  ggplot(aes(x = YearsSinceLastPromotion, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years Since Last Promotion")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p4 <- df_emp_ren %>%
  ggplot(aes(x = YearsWithCurrManager, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years With Current Manager")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

grid.arrange(p1, p2, p3, p4,  nrow = 2, ncol = 2)

```

### Employee Work Profile Demographics:  Insights
* Insights:
    1.  Employees with less tenure at the company are more likely to leave the company.
      + Hypothesis: This is a direct correlation with Years in current role as well as years with current manager.  
    2.  Employees with less time with their current manager then to have higher rates of attrition.
    + Hypothesis:  Employees may not be getting along with their current manager
    3.  Employees who have less years in their current role have higher rates of attrition.
    + Hypothesis:  Employees who have been placed in a role they do not like.
    4.  Employees with Low Job satisfaction tend to have high rates of attrition
    + Hypothesis:  Lower job satisfaction leads to higher attrition
    5.  Low Relationship status has higher attrition rates
    + Hypothesis:  Employees with Low Relationship status have higher attrition rates.
    6.  Employees with low employee satisfaction have higher rates of attrition.
    7.  Employees who are Sales Representatives have a high rate of attrition, as well as those who are in the sales department.
    8.  Employees who travel frequently have higher rates of attrition.
    9.  Employees how are newer to the company have a high rate of attrition.  

    
      
      
  


